Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Close #36
Add to paper draft

```{r}
library(hurricaneexposure)
```

@@

Current hurricane exposure assignments are often nonspecific and may lead to missclassification of exposure at the county-level [@Grabich2016]. Researchers have found heterogeneity in exposure assignments of counties in assessing multiple classification methods, both within and accross storms (i.e., storm dependent) [@Grabich2016]. Such missclassification can have important public health and economic implications.

Studies have begun to assess county-level hurricane exposure using large historical datasets [@Zandbergen2009] ... making the `hurricaneexposure` package a powerful tool in future exposure analysis.

Currently, it seems that most hurricane track mapping capabilities are restricted to geographical information system software. The introduction of this package expands the mapping potential of hurricane tracks and therefore the assessment of hurricane exposure and impacts to R.

#### Examples of using TS exposure estimates for multiple storms

[@Zandbergen2009] developed an exposure factor based on location (longitude and latitude), county shape, county size, and distance from the coast. The study assumed symmetrical activity of storm related factors,[@Zandbergen2009] which does not accurately represent storm characteristics once landfall has been made.[@Kruk2010, @Halverson2015]

@@

There are a number of limitations to using distance to assess exposure to tropical storms. Tropical storms vary in size and intensity, and a measure of distance from the storm track will not incorporate these differences and so could mis-classify exposure both in terms of generating false positives (counties close to the storm track of very mild or small-radius storms) and false negatives (e.g., during very large or intense storms).

@@

Currently, distance parameters involved with assessing risk of a particular storm have been rather arbitrary, contributing to the necessity in understanding how such parameters influence a county's exposure status. In assessing the risk of a given storm based on distance, recent studies have defined the distance from the storm track affected by a given storm differently. [@Czajkowski2011] assessed county-level risk and exposure based on a three-tierd definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary coutnies, and tertiary counties adjacent to secondary counties. Such a defenition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track [@Czajkowski2011]. Such a distance is slightly greater than that commonly used by public health departments (i.e., 100 km) [@Czajkowski2011]. In other cases, distance has been defined on the basis of maximum sustained wind speeds and their decay, with exposure defined as those counties exposed to wind speeds greater than or equal to tropical storm strength (39--73 miles per hour) [@Zandbergen2009] and may not provide an accurate respresentation of the communities exposed to a given storm (e.g., a large number of inland counties are likely underrepresented under this definition).

@@

Exposure classification based on rainfall has some advantages. For example, it allows the identification of exposed counties that are inland, rather than coastal, but that were affected by heavy rainfall during the storm. Often, storm-related deaths are associated with inland flooding. In fact, over half of the hurricane-related fatalities from 1970 to 1999 were a result of freshwater flooding, accounting for the vast majority of deaths in inland counties [@Rapport2000]. In another analysis, researchers found that 79\% of freshwater-drowning fatalities occured in inland counties,[@Czajkowski2011] further stressing the importance of rainfall in hurricane risk analysis. Storms can produce a lot of rain especially in certain topographies, like near mountains, so counties that are well inland can sometimes experience more extreme rain that other counties at similar distances from the storm's track.

In comparison to the storm track, the areas of extreme rain and extreme wind can vary. For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016]. In contrast, storm that interact with a downstream ridge tend to bring heavy rains to the right of the storm's center track [@Atallah2007].

@@

A variety of other wind speed models exist besides the one used here. In particular, there are options for wind speed models as far as ...

@@

@@

>>>>

```{r echo = FALSE}
library(hurricaneexposure)
```

@@

Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights). Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009]. Some studies suggest methods to determine exposure, including inland exposure, to tropical storms at the county-level over many years [@Zandbergen2009]. Here, we offer tools, through the open-source `hurricaneexposure` package, to determine county-level exposure to tropical storms using several different metrics (distance, rainfall, wind). In this paper, we describe how this software package assesses exposure to tropical storms at the county level, provide an analysis of the advantages and disadvantages of different strategies for assessing county-level exposure, and explore variation in exposure assessments based on using different exposure metrics.

Current hurricane exposure assignments are often nonspecific [GBA-- I found this wording a bit unclear] and may lead to missclassification of exposure at the county-level [@Grabich2016]. Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016]. Such missclassification can have important implications of assessments of the public health and economic impacts of tropical storms.

Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies]. Here, we offer methods to map and output historic exposure to tropical storms that does not require the use of proprietary software but instead uses a package written in the R statistical programming language, which is free and open-source.

#### Examples of estimating TS for multiple storms

When assessing exposure to a single tropical storm, a detailed analysis of ... can be performed for the storm. However, this level of detailed analysis is unreasonable when assessing exposure to all storms over an extended time period; further, such a long-term exposure assessment requires consistency in the data being used to determine levels of exposure to factors like wind and rain.

A few studies have sought to determine exposure to tropical storms over a  study period that includes multiple storms. For example, ... assessed the exposure of counties in Florida to hurricanes during the 2004 season [citation]. The largest county-level exposure study is likely that of @Zandbergen2009, which estimated exposure in US counties to all [US landfalling? or all storms?] Atlantic basin tropical storms between 1851 and 2003, using both distance and an exposure metric that incorporated distance and windspeed. They used this exposure evaluation to create maps of total exposure to tropical storms within US counties, as well as to explore associations between a county's long-term exposure to tropical storms and its location, distance from the coast, size, and shape [@Zandbergen2009].

@@

There are a number of limitations to using distance to assess exposure to tropical storms. Tropical storms vary in size and intensity, and a measure of distance from the storm track will not incorporate these differences and so could misclassify exposure both in terms of generating false positives (counties close to the storm track of very mild or small-radius storms) and false negatives (e.g., during very large or intense storms). In particular, the forces of a storm tend to be distributed around the center in a non-symmetrical way [citation], while distance-based exposure metrics that use buffers tend to use an equal buffer distance on each side of the storm track (e.g., [citations]).

Other distance-based exposure metrics have defined "exposure" to occur only when a storm's center track passes through a county's boundary (e.g., @Zandbergen2009, who describe this metric of exposure as a "hit" on the county). This way of measuring exposure can exclude nearby counties that suffered extreme conditions from the storm but were not directly on the hurricane's track, especially since hurricanes can have a width ([typical storm width; citation]) that is much larger than a typical county's width. Further, exposure assessed using this method has been found to be significantly [? double-check] associated with the size and shape of a county [@Zandbergen2009]. By contrast, assessing distance-based exposure based on measuring distance between the storm track and each county's center would be less susceptible to differences in county size and shape and so might be a more reliable metric of exposure than determining if a storm track ever crossed a county boundary. Further, assessing the minimum distance between county center and the storm track, as we do here, allows for a continuous, rather than binary, metric of distance-based exposure to a storm.

The @Zandbergen2009 study assumed symmetrical activity of storm related factors,which does not accurately represent storm characteristics once landfall has been made [@Kruk2010, @Halverson2015] [GBA-- I did not understand this thought. I agree that storms are asymmetric, but I don't think that's just after landfall. Is this something about characteristics of storms as they transform to extratropical systems? Let's discuss.]

@@

Currently, distance parameters involved with assessing risk of a particular storm have been rather arbitrary, contributing to the necessity in understanding how such parameters influence a county's exposure status. In assessing the risk of a given storm based on distance, recent studies have defined the distance from the storm track affected by a given storm differently. [@Czajkowski2011] assessed county-level risk and exposure based on a three-tierd definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary coutnies, and tertiary counties adjacent to secondary counties. Such a definition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track [@Czajkowski2011]. Such a distance is slightly greater than that commonly used by public health departments (i.e., 100 km) [@Czajkowski2011].

Another study used distance buffers of 30, 60, and 100 kilometers for different metrics of distance-based exposure [@Grabich2016] [check the paper-- was this distance to any part of a county, or to the center of the county?].

@@

Exposure classification based on rainfall has some advantages. For example, it allows the identification of exposed counties that are inland, rather than coastal, but that were affected by heavy rainfall during the storm. Often, storm-related deaths are associated with inland flooding. In fact, over half of the hurricane-related fatalities from 1970 to 1999 were a result of freshwater flooding, accounting for the vast majority of deaths in inland counties [@Rappaport2000]. In another analysis, researchers found that 79\% of freshwater-drowning fatalities occured in inland counties,[@Czajkowski2011] further stressing the importance of rainfall in hurricane risk analysis. Storms can produce a lot of rain especially in certain topographies, like near mountains, so counties that are well inland can sometimes experience more extreme rain that other counties at similar distances from the storm's track.

In comparison to the storm track, the areas of extreme rain and extreme wind can vary. For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point. The Halverson paper is great, but the Grabich paper I think might just be a secondary, rather than primary, reference for this point, so I think we could probably find a more direct reference]. Wind speeds are typically more severe to the right of the storm's track because here the counter-clockwise cyclonic [right word?] winds are moving in the same direction as the forward motion of the storm, creating an additive effect for total windspeeds [citation]. In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2007].

@@

A variety of other wind speed models exist besides the one used here. In particular, there are options for wind speed models as far as ... One study used a very simple model of constant winds up to a distance of 30 km from the storm track and a simple model of wind decay beyond that distance [@Zandbergen2009], based on a model developed by ... [see references for this paper]. Another study used wind data from a NOAA public database based on real-time hurricane wind analysis, rather than modeling windspeeds themselves [@Grabich2016] [see the Powell reference in this paper for more on that data].

@@

Exposure metrics based on wind speed may not provide an accurate representation of the communities exposed to a given storm, since, while wind speeds rapidly decay after landfall [citation], extreme rain can persist,for some  storms affecting a large number of inland counties, which would be underrepresented using an exposure metric the depends on windspeed.

Examples of studies that used windspeed as a sole criteria when determining tropical storm exposure include ... [citation]. In other studies, wind criteria have been combined with distance criteria, and with windspeed modeled based on maximum sustained windspeed estimates at each "best tracks" observation, combined with a simple model of decay in wind speed at distances further from the storm's center (wind speed equals the storm's maximum sustained windspeed up to 30 kilometers from the storm's center then follows a simple decay function beyond that distance) [@Zandbergen2009] [GBA-- this paper references another paper with more details on this decay model for windspeed-- the Gray and Klotzbach reference they have]. In this study, wind speed cutoffs of 40 mph, 75 mph, and 115 mph were used to determine if a county was exposed to a storm (tropical, hurricane-force, and intense hurricane-force, respectively) [@Zandbergen2009].

One study found that using a distance-based exposure metric that required the storm track to cross a county boundary assessed a number of counties as "unexposed" that suffered tropical storm- or hurricane-force winds from the storm and were assessed as "exposed" using a metric that combined distance and windspeed [@Zandbergen2009]; in fact, almost twice as many counties were assessed as having been exposed to at least one tropical storm between 1851 and 2003 when exposure was determined based on a combined wind-distance metric compared to a "hit" distance-based metric [@Zandbergen2009]. This suggests that some distance-based metrics might result in a number of false negatives for county-level storm exposure.

Studies that have used wind-based exposure metrics have used a variety of windspeed cutoff values to determine exposure. One study, for example, assigned a binary exposure value based on windspeed thresholds of 63 and 119 kilometers per hour, as well as a continuous exposure metric based on the exact value of the windspeed in the county during the storm and a categorical exposure metric based on the Saffir-Simpson storm severity categories [@Grabich2016].

#### Other ways of assessing tropical storm exposure

One study used FEMA disaster declarations as a binary indicator of tropical storm exposure [@Grabich2016]. They found that this metric tended to assess a lot more counties as "exposed" to a storm than distance- or wind-based metrics [@Grabich2016], although they only compared the two exposure metrics as applied to four storms (Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in 2004).

@@

#### Table including various exposure papers and the parameters used

### 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.tex
Add to paper draft

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Assessing exposure to Atlantic Basin tropical storms in United States counties},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(hurricaneexposure)}
\end{Highlighting}
\end{Shaded}

@@

Current hurricane exposure assignments are often nonspecific and may
lead to missclassification of exposure at the county-level (Grabich et
al. 2016). Researchers have found heterogeneity in exposure assignments
of counties in assessing multiple classification methods, both within
and accross storms (i.e., storm dependent) (Grabich et al. 2016). Such
missclassification can have important public health and economic
implications.

Studies have begun to assess county-level hurricane exposure using large
historical datasets (Zandbergen 2009) \ldots{} making the
\texttt{hurricaneexposure} package a powerful tool in future exposure
analysis.

Currently, it seems that most hurricane track mapping capabilities are
restricted to geographical information system software. The introduction
of this package expands the mapping potential of hurricane tracks and
therefore the assessment of hurricane exposure and impacts to R.

\paragraph{Examples of using TS exposure estimates for multiple
storms}\label{examples-of-using-ts-exposure-estimates-for-multiple-storms}

(Zandbergen 2009) developed an exposure factor based on location
(longitude and latitude), county shape, county size, and distance from
the coast. The study assumed symmetrical activity of storm related
factors,(Zandbergen 2009) which does not accurately represent storm
characteristics once landfall has been made.(Kruk et al. 2010, Halverson
(2015))

@@

There are a number of limitations to using distance to assess exposure
to tropical storms. Tropical storms vary in size and intensity, and a
measure of distance from the storm track will not incorporate these
differences and so could mis-classify exposure both in terms of
generating false positives (counties close to the storm track of very
mild or small-radius storms) and false negatives (e.g., during very
large or intense storms).

@@

Currently, distance parameters involved with assessing risk of a
particular storm have been rather arbitrary, contributing to the
necessity in understanding how such parameters influence a county's
exposure status. In assessing the risk of a given storm based on
distance, recent studies have defined the distance from the storm track
affected by a given storm differently. (Czajkowski, Simmons, and Sutter
2011) assessed county-level risk and exposure based on a three-tierd
definition, with primary counties being those closest to the storm track
on either side, secondary counties being adjacent to primary coutnies,
and tertiary counties adjacent to secondary counties. Such a defenition
resulted in an exposure defenition based on an average distance radius
of 120 km on either side of the storm track (Czajkowski, Simmons, and
Sutter 2011). Such a distance is slightly greater than that commonly
used by public health departments (i.e., 100 km) (Czajkowski, Simmons,
and Sutter 2011). In other cases, distance has been defined on the basis
of maximum sustained wind speeds and their decay, with exposure defined
as those counties exposed to wind speeds greater than or equal to
tropical storm strength (39--73 miles per hour) (Zandbergen 2009) and
may not provide an accurate respresentation of the communities exposed
to a given storm (e.g., a large number of inland counties are likely
underrepresented under this definition).

@@

In comparison to the storm track, the areas of extreme rain and extreme
wind can vary. For example, storms undergoing an extratropical
transition can bring heavy rains to the left of the storm's center
track, while extreme winds are more common to the right of the track
(Halverson 2015, Grabich et al. (2016)). In contrast, storm that
interact with a downstream ridge tend to bring heavy rains to the right
of the storm's center track (Atallah, Bosart, and Aiyyer 2007).

@@

A variety of other wind speed models exist besides the one used here. In
particular, there are options for wind speed models as far as \ldots{}

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Assessing exposure to Atlantic Basin tropical storms in United States counties},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

@@

Many outcomes of interest are available at county-level aggregations
(e.g., counts of health outcomes: Grabich et al. (2016): premature
births and low birth weights). Further, decisions and policies to
prepare for and respond to storms are often undertaken at the county
level (Zandbergen 2009). Some studies suggest methods to determine
exposure, including inland exposure, to tropical storms at the
county-level over many years (Zandbergen 2009). Here, we offer tools,
through the open-source \texttt{hurricaneexposure} package, to determine
county-level exposure to tropical storms using several different metrics
(distance, rainfall, wind). In this paper, we describe how this software
package assesses exposure to tropical storms at the county level,
provide an analysis of the advantages and disadvantages of different
strategies for assessing county-level exposure, and explore variation in
exposure assessments based on using different exposure metrics.

Current hurricane exposure assignments are often nonspecific {[}GBA-- I
found this wording a bit unclear{]} and may lead to missclassification
of exposure at the county-level (Grabich et al. 2016). Researchers have
found heterogeneity in exposure assignments of counties, both within and
between storms, when different metrics of exposure are used (Grabich et
al. 2016). Such missclassification can have important implications of
assessments of the public health and economic impacts of tropical
storms.

Many of the studies that have assessed exposure to tropical storms in
the US have used geographical information system software (e.g., ArcGIS)
{[}example studies{]}. Here, we offer methods to map and output historic
exposure to tropical storms that does not require the use of proprietary
software but instead uses a package written in the R statistical
programming language, which is free and open-source.

\paragraph{Examples of estimating TS for multiple
storms}\label{examples-of-estimating-ts-for-multiple-storms}

When assessing exposure to a single tropical storm, a detailed analysis
of \ldots{} can be performed for the storm. However, this level of
detailed analysis is unreasonable when assessing exposure to all storms
over an extended time period; further, such a long-term exposure
assessment requires consistency in the data being used to determine
levels of exposure to factors like wind and rain.

A few studies have sought to determine exposure to tropical storms over
a study period that includes multiple storms. For example, \ldots{}
assessed the exposure of counties in Florida to hurricanes during the
2004 season {[}citation{]}. The largest county-level exposure study is
likely that of Zandbergen (2009), which estimated exposure in US
counties to all {[}US landfalling? or all storms?{]} Atlantic basin
tropical storms between 1851 and 2003, using both distance and an
exposure metric that incorporated distance and windspeed. They used this
exposure evaluation to create maps of total exposure to tropical storms
within US counties, as well as to explore associations between a
county's long-term exposure to tropical storms and its location,
distance from the coast, size, and shape (Zandbergen 2009).

@@

There are a number of limitations to using distance to assess exposure
to tropical storms. Tropical storms vary in size and intensity, and a
measure of distance from the storm track will not incorporate these
differences and so could misclassify exposure both in terms of
generating false positives (counties close to the storm track of very
mild or small-radius storms) and false negatives (e.g., during very
large or intense storms). In particular, the forces of a storm tend to
be distributed around the center in a non-symmetrical way
{[}citation{]}, while distance-based exposure metrics that use buffers
tend to use an equal buffer distance on each side of the storm track
(e.g., {[}citations{]}).

Other distance-based exposure metrics have defined ``exposure'' to occur
only when a storm's center track passes through a county's boundary
(e.g., Zandbergen (2009), who describe this metric of exposure as a
``hit'' on the county). This way of measuring exposure can exclude
nearby counties that suffered extreme conditions from the storm but were
not directly on the hurricane's track, especially since hurricanes can
have a width ({[}typical storm width; citation{]}) that is much larger
than a typical county's width. Further, exposure assessed using this
method has been found to be significantly {[}? double-check{]}
associated with the size and shape of a county (Zandbergen 2009). By
contrast, assessing distance-based exposure based on measuring distance
between the storm track and each county's center would be less
susceptible to differences in county size and shape and so might be a
more reliable metric of exposure than determining if a storm track ever
crossed a county boundary. Further, assessing the minimum distance
between county center and the storm track, as we do here, allows for a
continuous, rather than binary, metric of distance-based exposure to a
storm.

The Zandbergen (2009) study assumed symmetrical activity of storm
related factors,which does not accurately represent storm
characteristics once landfall has been made (Kruk et al. 2010, Halverson
(2015)) {[}GBA-- I did not understand this thought. I agree that storms
are asymmetric, but I don't think that's just after landfall. Is this
something about characteristics of storms as they transform to
extratropical systems? Let's discuss.{]}

@@

Currently, distance parameters involved with assessing risk of a
particular storm have been rather arbitrary, contributing to the
necessity in understanding how such parameters influence a county's
exposure status. In assessing the risk of a given storm based on
distance, recent studies have defined the distance from the storm track
affected by a given storm differently. (Czajkowski, Simmons, and Sutter
2011) assessed county-level risk and exposure based on a three-tierd
definition, with primary counties being those closest to the storm track
on either side, secondary counties being adjacent to primary coutnies,
and tertiary counties adjacent to secondary counties. Such a definition
resulted in an exposure defenition based on an average distance radius
of 120 km on either side of the storm track (Czajkowski, Simmons, and
Sutter 2011). Such a distance is slightly greater than that commonly
used by public health departments (i.e., 100 km) (Czajkowski, Simmons,
and Sutter 2011).

Another study used distance buffers of 30, 60, and 100 kilometers for
different metrics of distance-based exposure (Grabich et al. 2016)
{[}check the paper-- was this distance to any part of a county, or to
the center of the county?{]}.

@@

In comparison to the storm track, the areas of extreme rain and extreme
wind can vary. For example, storms undergoing an extratropical
transition can bring heavy rains to the left of the storm's center
track, while extreme winds are more common to the right of the track
(Halverson 2015, Grabich et al. (2016) GBA-- I think it would be better
to just use meteorological papers as references for this point. The
Halverson paper is great, but the Grabich paper I think might just be a
secondary, rather than primary, reference for this point, so I think we
could probably find a more direct reference). Wind speeds are typically
more severe to the right of the storm's track because here the
counter-clockwise cyclonic {[}right word?{]} winds are moving in the
same direction as the forward motion of the storm, creating an additive
effect for total windspeeds {[}citation{]}. In contrast, rains can be
heaviest to the left of the storm's track, especially in cases when the
storm interacts with a downstream ridge (Atallah, Bosart, and Aiyyer
2007).

@@

A variety of other wind speed models exist besides the one used here. In
particular, there are options for wind speed models as far as \ldots{}
One study used a very simple model of constant winds up to a distance of
30 km from the storm track and a simple model of wind decay beyond that
distance (Zandbergen 2009), based on a model developed by \ldots{}
{[}see references for this paper{]}. Another study used wind data from a
NOAA public database based on real-time hurricane wind analysis, rather
than modeling windspeeds themselves (Grabich et al. 2016) {[}see the
Powell reference in this paper for more on that data{]}.

@@

Exposure metrics based on wind speed may not provide an accurate
representation of the communities exposed to a given storm, since, while
wind speeds rapidly decay after landfall {[}citation{]}, extreme rain
can persist,for some storms affecting a large number of inland counties,
which would be underrepresented using an exposure metric the depends on
windspeed.

Examples of studies that used windspeed as a sole criteria when
determining tropical storm exposure include \ldots{} {[}citation{]}. In
other studies, wind criteria have been combined with distance criteria,
and with windspeed modeled based on maximum sustained windspeed
estimates at each ``best tracks'' observation, combined with a simple
model of decay in wind speed at distances further from the storm's
center (wind speed equals the storm's maximum sustained windspeed up to
30 kilometers from the storm's center then follows a simple decay
function beyond that distance) (Zandbergen 2009) {[}GBA-- this paper
references another paper with more details on this decay model for
windspeed-- the Gray and Klotzbach reference they have{]}. In this
study, wind speed cutoffs of 40 mph, 75 mph, and 115 mph were used to
determine if a county was exposed to a storm (tropical, hurricane-force,
and intense hurricane-force, respectively) (Zandbergen 2009).

One study found that using a distance-based exposure metric that
required the storm track to cross a county boundary assessed a number of
counties as ``unexposed'' that suffered tropical storm- or
hurricane-force winds from the storm and were assessed as ``exposed''
using a metric that combined distance and windspeed (Zandbergen 2009);
in fact, almost twice as many counties were assessed as having been
exposed to at least one tropical storm between 1851 and 2003 when
exposure was determined based on a combined wind-distance metric
compared to a ``hit'' distance-based metric (Zandbergen 2009). This
suggests that some distance-based metrics might result in a number of
false negatives for county-level storm exposure.

Studies that have used wind-based exposure metrics have used a variety
of windspeed cutoff values to determine exposure. One study, for
example, assigned a binary exposure value based on windspeed thresholds
of 63 and 119 kilometers per hour, as well as a continuous exposure
metric based on the exact value of the windspeed in the county during
the storm and a categorical exposure metric based on the Saffir-Simpson
storm severity categories (Grabich et al. 2016).

\paragraph{Other ways of assessing tropical storm
exposure}\label{other-ways-of-assessing-tropical-storm-exposure}

One study used FEMA disaster declarations as a binary indicator of
tropical storm exposure (Grabich et al. 2016). They found that this
metric tended to assess a lot more counties as ``exposed'' to a storm
than distance- or wind-based metrics (Grabich et al. 2016), although
they only compared the two exposure metrics as applied to four storms
(Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in 2004).
